Efficiently Handling Guards in the low level design of RPython's tracing JIT

### 2012-07-17 talk/vmil2012/paper.tex
expand notes somewhat

The RPython language and the PyPy Project were started in 2002 with the goal of
creating a python interpreter written in a High level language, allowing easy
language experimentation and extension. PyPy is now a fully compatible
alternative implementation of the Python language\bivab{mention speed}. The
Implementation takes advantage of the language features provided by RPython
such as the provided tracing just-in-time compiler described below.

@@

* High level handling of resumedata
   * trade-off fast tracing v/s memory usage
   * creation in the frontend
   * optimization
   * compression
   * interaction with optimization
   * tracing and attaching bridges and throwing away resume data
   * compiling bridges

>>>>

The RPython language and the PyPy Project were started in 2002 with the goal of
creating a Python interpreter written in a high level language, allowing easy
language experimentation and extension. PyPy is now a fully compatible
alternative implementation of the Python language\bivab{mention speed}. The
Implementation takes advantage of the language features provided by RPython
such as the provided tracing just-in-time compiler described below.

@@

* High level handling of resumedata

- traces follow the execution path during tracing, other path not compiled at first
- points of possible divergence from that path are guards
- since path can later diverge, at the guards it must be possible to re-build interpreter state in the form of interpreter stack frames
- tracing does inlining, therefore a guard must contain information to build a whole stack of frames
- optimization rewrites traces, including removal of guards

- frames consist of a PC and local variables
- rebuild frame by taking local SSA variables in the trace and mapping them to variables in the frame

two forces:
- there are lots of guards, therefore the information must be stored in a compact way in the end
- tracing must be fast

compression approaches:
- use fact that outer frames don't change in the part of the trace that is in the inner frame
- compact bit-representation for constants/ssa vars

interaction with optimization
- guard coalescing
- virtuals

   * tracing and attaching bridges and throwing away resume data
   * compiling bridges
